\propertitle

### 2020-03-16 replies1_ncomp.tex
update

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2020-03-16T19:55:02+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourtrue% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimfalse
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\firstdraft}{16 March 2020}
\newcommand*{\firstpublished}{16 March 2020}
\newcommand*{\updated}{\ifarxiv***\else\today\fi}
\newcommand*{\propertitle}{Sketch of replies to N. Comp. reviews%\\{\large ***}%
}% title uses LARGE; set Large for smaller
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{}
\newcommand*{\pdfauthor}{}
\newcommand*{\headauthor}{}
\newcommand*{\reporthead}{}%\ifarxiv\else Open Science Framework \href{https://doi.org/10.31219/osf.io/***}{\textsc{doi}:10.31219/osf.io/***}\fi}% Report number

@@

\item As $N$ increases, the two peaks we infer do become more pronounced,
  as suggested by our Fig.~1. So our inference is not really in
  contradiction with Amari \etal. The results of that paper hold for the
  $N\to\infty$, and no explicit form for the limit-remainder is given. It
  is therefore unclear whether at $N=10\,000$ we should already observe a
  delta. Our findings~\ref{item:maxmomentcheck}, above, suggest that we
  should not. Limit-reminders such as \enquote{$\Ord(1/N^{k})$} as given in
  Amari \etal\ do not help unless we know their precise form. As an extreme
  example, $\frac{10^{10}}{N}$ is $\Ord(1/N)$, and in this case the $1/N$
  term cannot really be neglected unless $N \gg 10^{10}$.

@@

In any case, please note that we avoid speaking about
\enquote{significance}, also as recommended by the American Statistical
Association in their recent official statements \citep{asa2016}[see
especially \sect~2]{asa2019}. When comparing hypotheses $M_{i}$ we simply
want their post-data probabilities $\p(M_{i} \| D)$, with their continuum
of cases. Other researchers are of course free to make dichotomies if they
like; but we personally agree with the American Statistical Association
that any threshold is unnecessary and arbitrary. If a \emph{choice} among
the hypotheses has to be made for some purpose (say, hardware algorithm
implementation in some automated device), such choice requires not only the
probabilities $\p(M_{i}\|D)$, but also the utility/loss matrix for the
various choices conditional on the various hypotheses
\citep{kadaneetal1980b}. The choice is then made by maximization of the
expected utility, a procedure dictated by some simple rational desiderata
\citep{fishburn1981,bernardoetal1994_r2000}[\chap~13]{jaynes1994_r2003}.
Some utility matrices can even lead to choosing the hypothesis having
\emph{lower} post-data probability; this frequently happens in
medical decision problems, see \eg\ \textcite{soxetal1988_r2013}.

@@

The main question is what we mean by \enquote{model}. There are two main
possibilities: model as a family, or model as a specific distribution.

@@

\textbf{(A)} If by \enquote{model} we mean a family of probability
distributions, \eg\ (apart from normalization)
\[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
    \mu\in\RR, \sigma \in\opop{0,\infty}}\;,\]
then model comparison is ill-defined. % Each distribution in the family
% assigns a different probability, and it is not clear which should be
% chosen or how they should be combined and why.
It may happen that the optimal distribution in family $\mathcal{F}_{1}$
assigns more probability to the observed data than the optimal distribution
in family $\mathcal{F}_{2}$, and yet $\mathcal{F}_{1}$ assigns \enquote{on
  average} lower probabilities to the data than $\mathcal{F}_{2}$. It is
not clear how the two families should be compared in this case, or what
$\p(D \| \mathcal{F}_{i})$ is. A hypothesis is ill-defined, or at least
useless, if it does not allow us to assign a probability to the data. Such
a probability can be obtained if we specify a pre-data probability
distribution over the distributions in the family, \eg\
$\pf(\mu,\sigma \| I)\,\di\mu\,\di\sigma$ in the example above.
\enquote{Model} must therefore be redefined as a family of distributions
\emph{and} a pre-data distribution over such family. This redefinition has
several consequences:

First, two models are different if they have the same family but different
distributions over it. In fact they lead to different probabilities on the
data space. Two such models are also mutually exclusive,
$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use the one family
distribution \emph{and} the other family distribution at the same time.
It's either-or. This consequence makes sense from a nonparametric
perspective, where a parametric model (in the redefined sense) is just a
delta-like nonparametric distribution, concentrated on a lower-dimensional
manifold (the family above) of the full space of distributions
\citep[p.~761]{draperetal1996}.

Second, the \enquote{complexity} of a model cannot be judged from the
dimensionality of its family. Model $M_{1}$ can have a family distribution
concentrated on a very small region, with respect to model $M_{2}$ having
the same family. $M_{1}$ is thus effectively using fewer parameters than
$M_{2}$. Correspondingly the \enquote{typical set} of data to which $M_{1}$
assigns appreciable probability is smaller than the typical set of $M_{2}$:
model $M_{1}$ is less \enquote{powerful}. The effective dimensionality or
complexity of a model is thus best defined by the volume of data that is
typical under that model. Most important, it may then happen that a model
is more powerful than another even if it is defined on a lower-dimensional
family.

Third, \emph{the evidence $\p(D \| M_{i})$ of a model automatically takes
  care of its complexity}. This happens because the probability over the
data space is normalized. If $M_{1}$ has a typical-data volume larger than
$M_{2}$'s, then it must also assign lower probability to each data point in
that volume, to ensure normalization. As a result, if the observed data
$D^{*}$ lie in the typical volumes of both models, then
$\pf(D^{*} \| M_{1})< \pf(D^{*} \| M_{2})$: \emph{the evidence is
  automatically penalizing the larger model}. There is a continuous
trade-off between how typical the data is for each model and each model's
power. This is the so-called automatic Ockham razor of Bayes's theorem,
extensively explained by MacKay
\citey{mackay1992}[also][\sect~28.1]{mackay1995_r2003}. And this is the
reason why we are personally satisfied with simply comparing the evidence
for any two models.

In our calculations, however, we are using a zeroth-order Laplace
approximation for calculating the evidence in \sect~4. A first-order
approximation (containing the width of the distribution, influenced by the
typical-set volume above) would correct for the model's power. We believe
that the correction would nevertheless be the same in the calculation based
on the larger population, \eqns~(11), and on the sample, \eqns~(12),
because the data space is the same. Hence the difference in the two
approaches -- our main message -- should persist. But it would be better to
check this.

@@

But the property of having a minimal sufficient statistic is still
well-defined for a model in this sense, in an exchangeable context. The
model above has $\set[\big]{n, \widebar{x}, \widebar{x^{2}}}$ (or any
equivalent set) as minimal sufficient statistics, where the bar denotes the
sample average. Thus two models having different sufficient statistics can
be compared without reference to any family. Interpreted in this sense our
calculations of \sect~4 are correct.

@@

The question, in summary, is what exactly is being compared, whether
families (with accompanying distribution) or single distributions. The
literature is not at all clear. But, in either case, our calculations show
that the approaches based on a population and a sample thereof can lead to
different conclusions -- in accordance with the non-preservation of
sufficiency under sampling -- and they provide a way to bypass this
problem.

>>>>

\pdfoutput=1
%% Author: PGL  Porta Mana
%% Created: 2015-05-01T20:53:34+0200
%% Last-Updated: 2020-03-16T20:21:43+0100
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\newif\ifarxiv
\arxivfalse
\ifarxiv\pdfmapfile{+classico.map}\fi
\newif\ifafour
\afourtrue% true = A4, false = A5
\newif\iftypodisclaim % typographical disclaim on the side
\typodisclaimfalse
\newcommand*{\memfontfamily}{zplx}
\newcommand*{\memfontpack}{newpxtext}
\documentclass[\ifafour a4paper,12pt,\else a5paper,10pt,\fi%extrafontsizes,%
onecolumn,oneside,article,%french,italian,german,swedish,latin,
british%
]{memoir}
\newcommand*{\firstdraft}{16 March 2020}
\newcommand*{\firstpublished}{16 March 2020}
\newcommand*{\updated}{\ifarxiv***\else\today\fi}
\newcommand*{\propertitle}{Sketch of replies to N. Comp. reviews%\\{\large ***}%
}% title uses LARGE; set Large for smaller
\newcommand*{\pdftitle}{\propertitle}
\newcommand*{\headtitle}{}
\newcommand*{\pdfauthor}{}
\newcommand*{\headauthor}{}
\newcommand*{\reporthead}{}%\ifarxiv\else Open Science Framework \href{https://doi.org/10.31219/osf.io/***}{\textsc{doi}:10.31219/osf.io/***}\fi}% Report number

@@

\item As $N$ increases, the two peaks we infer do become more pronounced,
  as suggested by our Fig.~1. So our inference is not really in
  contradiction with Amari \etal. The results of that paper hold for the
  $N\to\infty$ limit, and no explicit form for the limit-remainder is given
  there. Limit-reminders such as \enquote{$\Ord(1/N^{k})$} as given in
  Amari \etal\ do not help unless we know their precise form. As an extreme
  example, $\frac{10^{10}}{N}$ is $\Ord(1/N)$, and in this case the $1/N$
  term cannot really be neglected unless $N \gg 10^{10}$. It is therefore
  unclear whether at $N=10\,000$ we should already observe a delta. Our
  findings~\ref{item:maxmomentcheck}, above, suggest that we should not.

@@

In any case, please note that we avoid speaking about
\enquote{significance}, also as recommended by the American Statistical
Association in their recent official statements \citep{asa2016}[see
especially \sect~2]{asa2019}. When comparing hypotheses $M_{i}$ we want
their full post-data probabilities $\p(M_{i} \| D)$, with their continuum
of cases. Other researchers are of course free to make dichotomies if they
like; but we personally agree with the American Statistical Association
that any threshold is unnecessary and arbitrary. If a \emph{choice} among
the hypotheses has to be made for some purpose (say, hardware algorithm
implementation in some automated device), such choice requires not only the
probabilities $\p(M_{i}\|D)$, but also the utility/loss matrix for the
various choices conditional on the various hypotheses
\citep{kadaneetal1980b}. The choice is then made by maximization of the
expected utility, a procedure dictated by some simple rational desiderata
\citep{fishburn1981,bernardoetal1994_r2000}[\chap~13]{jaynes1994_r2003}.
Some utility matrices can even lead to choosing the hypothesis having
\emph{lower} post-data probability; this frequently happens in medical
decision problems, see \eg\ \textcite{soxetal1988_r2013}.

@@

The main question, on the other hand, is what we mean by \enquote{model}.
There are two main possibilities\ibid: model as a family, or model as a specific
distribution.

@@

\textbf{(A)} If by \enquote{model} we mean a family of probability
distributions, \eg\ (apart from normalization)
\[\conds*{\exp\biggl[-\frac{(x-\mu)^{2}}{2\sigma^{2}}\biggr]\,\di x}{
    \mu\in\RR, \sigma \in \RR_{>0}}\;,\]
then model comparison is ill-defined. % Each distribution in the family
% assigns a different probability, and it is not clear which should be
% chosen or how they should be combined and why.
It may happen that the optimal distribution in family $\mathcal{F}_{1}$
assigns more probability to the observed data than the optimal distribution
in family $\mathcal{F}_{2}$, and yet $\mathcal{F}_{1}$ assigns \enquote{on
  average} lower probabilities to the data than $\mathcal{F}_{2}$. It is
not clear how the two families should be compared in this case, or what
$\p(D \| \mathcal{F}_{i})$ is. A hypothesis is ill-defined, or at least
useless, if it does not allow us to assign a probability to the data. Such
a probability can be obtained if we specify a pre-data probability
distribution over the distributions in the family, \eg\
$\pf(\mu,\sigma \| I)\,\di\mu\,\di\sigma$ in the example above.
\enquote{Model} must therefore be redefined as a family of distributions
\emph{and} a pre-data distribution over such family. This redefinition has
several consequences:

First, two models are different if they have the same family but different
distributions over it. In fact they lead to different probabilities on the
data space. Two such models are also mutually exclusive,
$\p(M_{1} \land M_{2} \|I)=0$, because we cannot use $M_{1}$'s family
distribution \emph{and} $M_{2}$'s family distribution at the same time.
It's either-or. This consequence makes sense from a nonparametric
perspective, where a parametric model is just a delta-like nonparametric
distribution, concentrated on a lower-dimensional manifold (the family
above) of the full space of distributions \citep[p.~761]{draperetal1996}.
The notion of nested models becomes unimportant, from this perspective.

Second, the \enquote{complexity} of a model cannot be judged from the
dimensionality of its family. Models $M_{1}$ and $M_{2}$ can have the same
family, but $M_{1}$ can have a family distribution concentrated on a very
small region, with respect to that of $M_{2}$. Model $M_{1}$ is thus
effectively using fewer parameters than $M_{2}$. Correspondingly the
\enquote{typical set} of data to which $M_{1}$ assigns appreciable
probability is smaller than the typical set of $M_{2}$: model $M_{1}$ is
less \enquote{powerful}. The effective dimensionality, complexity, or power
of a model is thus best defined by the volume of \emph{data space} that is
typical under that model. Most important, it may then happen that a model
is more powerful than another even if it is based on a lower-dimensional
family.

Third, \emph{the evidence $\p(D \| M_{i})$ of a model automatically takes
  care of its complexity}. This happens because the probability over the
data space is normalized. If $M_{2}$ has a typical-data volume larger than
$M_{1}$'s, then it must also assign lower probability to each volume
element in that volume, to ensure normalization. As a result, if the
observed data $D^{*}$ lie in the typical volumes of both models, then
$\pf(D^{*} \| M_{1}) > \pf(D^{*} \| M_{2})$: \emph{the evidence is
  automatically penalizing the larger model}. There is a continuous
trade-off between how typical the observed data are for each model and each
model's power. This is the so-called automatic Ockham razor of Bayes's
theorem, extensively explained by MacKay
\citey{mackay1992}[also][\sect~28.1]{mackay1995_r2003}. And this is the
reason why we are personally satisfied with simply comparing the evidence
for any two models.

In our calculations, however, we are using a zeroth-order Laplace
approximation for calculating the evidence in \sect~4. A first-order
approximation (containing the width of the distribution, influenced by the
typical-set volume above) would correct for the model's power. We believe
that the correction would nevertheless be the same in the calculation based
on the larger population, \eqns~(11), and on the sample, \eqns~(12),
because the data space is the same in both. Hence the difference in the two
approaches -- our main message -- should persist. But it would be better to
check this.

@@

Yet, the property of having a minimal sufficient statistic is still
well-defined for a model in this sense, in an exchangeable context. The
model above has $\set[\big]{n, \widebar{x}, \widebar{x^{2}}}$ (or any
equivalent set) as minimal sufficient statistics, where the bar denotes the
sample average. Thus two models having different sufficient statistics can
be compared without reference to any family. Interpreted in this sense our
calculations of \sect~4 are correct.

@@

The question, in summary, is what exactly is being compared, whether
families (and what their accompanying distributions are) or single
distributions. The literature is not at all clear. But, in either case, our
calculations show that the approaches based on a population and a sample
thereof can lead to different conclusions -- in accordance with the
non-preservation of sufficiency under sampling -- and they provide a way to
bypass this problem.
